Type Error Telemetry at Scale

### 2023-09-26 programming23/paper.tex
try FS error density, uhoh its an even split 1/3 each

  \paragraph{Importance}
  %% Why does this work matter?
  Luau is the first large-scale combination of gradual types and semantic
  subtyping.
  Beyond the immediate benefits to Luau,
  our findings about types and type errors have implications
  for adoption and ergonomics in other gradual languages.
  %% mention Elixir, Verse?
  FILL specifics.
  Our telemetry design is of broad interest, as it gathers useful
  data without exposing personal information.

@@

%% TODO migratory

@@

% TODO
% \begin{table}
%   \caption{Modes and analyses}
%   \label{t:mode-analyze}
% 
%   Mode: nocheck ...
%   Analysis: nocheck ... forced strict
% 
% \end{table}

@@

\begin{figure}[t]\centering
  % TODO try diff vs first record instead of previous?

  \mnocheck{}
  \includegraphics[width=\columnwidth]{img/error-count-nocheck-row--te-density-diff.pdf}
  \medskip
  \mnonstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-nonstrict-row--te-density-diff.pdf}
  \medskip
  \mstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-strict-row--te-density-diff.pdf}
  \caption{Changes over time (seconds) to type error density (total errors / lines in project).}
  \label{f:error-density}
\end{figure}

\FILL{} what does the FS plot look like? is that a better nocheck-vs-rest comparison??

@@

\begin{figure}[t]\centering
  % TODO use bars instead, percent only, don't care about counts
  %% TODO old version had +-20% add/drop and a bigger % of keeps. What changed?
  \begin{tabular}{lr@{}rr@{}rr@{}r}
    & \zerowidth{Add} & & \zerowidth{Keep} & & \zerowidth{Drop} \\\midrule
    \mnocheck{} & 48378 & [\pct{46.83}] & 9440 & [\pct{9.14}] & 45479 & [\pct{44.03}] \\
    \mnonstrict{} & 19491 & [\pct{39.63}] & 9567 & [\pct{19.45}] & 20121 & [\pct{40.91}] \\
    \mstrict{} & 733 & [\pct{39.18}] & 368 & [\pct{19.67}] & 770 & [\pct{41.15}] \\
  \end{tabular}
  \caption{How often do edits increase, maintain, or decrease the number of type errors?}
  \label{f:error-changes}
\end{figure}

@@

\section{Reflections and Lessons Learned}
\label{s:conclusion}

@@

In future work, it would be interesting to build statistical models of
the ``average'' programmer using each typing mode and compare their error
rates.
It would be possible to start modeling with our \mnocheck{} and \mnonstrict{}
data, but we have too few \mstrict{} records at the moment due to our uniform
sampling rate.
More data is needed  for a robust model of \mstrict{} programmers.

%% PS downloading from Kibana was a HUGE pain and source of errors. Awful!

>>>>

  \paragraph{Importance}
  %% Why does this work matter?
  Luau is the first large-scale combination of gradual types and semantic
  subtyping.
  Beyond the immediate benefits to Luau,
  our findings about types and type errors have implications
  for adoption and ergonomics in other gradual languages.
  %% mention Elixir, Verse?
  FILL specifics.
  %% TODO remove semantic subtyping
  Our telemetry design is of broad interest, as it gathers useful
  data without exposing personal information.

@@

%% TODO Luau is migratory, clearer about this

@@

% TODO quick reference table
% \begin{table}
%   \caption{Modes and analyses}
%   \label{t:mode-analyze}
% 
%   Mode: nocheck ...
%   Analysis: nocheck ... forced strict
% 
% \end{table}

@@

\begin{figure}[t]\centering
  %% FS plots look similar
  %% - much higher densities, same basic story
  %% - nocheck indeed the highest > NS > S

  \mnocheck{}
  \includegraphics[width=\columnwidth]{img/error-count-nocheck-row--te-density-diff.pdf}
  \medskip
  \mnonstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-nonstrict-row--te-density-diff.pdf}
  \medskip
  \mstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-strict-row--te-density-diff.pdf}
  \caption{Changes in type error density (errors in module / lines in module) over time (seconds).}
  \label{f:error-density}
\end{figure}

@@

\begin{figure}[t]\centering
  % TODO use bars instead, percent only, don't care about counts

  % TE: nocheck & 48378 & [\pct{46.83}] & 9440 & [\pct{9.14}] & 45479 & [\pct{44.03}] \\
  % TE: nonstrict & 19491 & [\pct{39.63}] & 9567 & [\pct{19.45}] & 20121 & [\pct{40.91}] \\
  % TE: strict & 733 & [\pct{39.18}] & 368 & [\pct{19.67}] & 770 & [\pct{41.15}] \\
  %
  % TE mod: nocheck & 20085 & [\pct{49.83}] & 906 & [\pct{2.25}] & 19317 & [\pct{47.92}] \\
  % TE mod: nonstrict & 13030 & [\pct{40.22}] & 5513 & [\pct{17.02}] & 13852 & [\pct{42.76}] \\
  % TE mod: strict & 419 & [\pct{39.31}] & 230 & [\pct{21.58}] & 417 & [\pct{39.12}] \\
  %
  % FS: nocheck & 270763 & [\pct{36.99}] & 238696 & [\pct{32.61}] & 222588 & [\pct{30.41}] \\
  % FS: nonstrict & 35330 & [\pct{37.28}] & 29483 & [\pct{31.11}] & 29955 & [\pct{31.61}] \\
  % FS: strict & 574 & [\pct{33.71}] & 561 & [\pct{32.94}] & 568 & [\pct{33.35}] \\
  %
  % FS mod: nocheck & 259865 & [\pct{39.09}] & 175643 & [\pct{26.42}] & 229271 & [\pct{34.49}] \\
  % FS mod: nonstrict & 33988 & [\pct{39.74}] & 20738 & [\pct{24.25}] & 30798 & [\pct{36.01}] \\
  % FS mod: strict & 488 & [\pct{36.53}] & 344 & [\pct{25.75}] & 504 & [\pct{37.72}] \\
  \begin{tabular}{lr@{}rr@{}rr@{}r}
    & \zerowidth{Add} & & \zerowidth{Keep} & & \zerowidth{Drop} \\\midrule
    \mnocheck{}   & 20085 & [\pct{49.83}] & 906 & [\pct{2.25}] & 19317 & [\pct{47.92}] \\
    \mnonstrict{} & 13030 & [\pct{40.22}] & 5513 & [\pct{17.02}] & 13852 & [\pct{42.76}] \\
    \mstrict{}    & 419 & [\pct{39.31}] & 230 & [\pct{21.58}] & 417 & [\pct{39.12}] \\
  \end{tabular}
  \caption{How often do edits increase, maintain, or decrease the number of errors?}
  \label{f:error-changes}
\end{figure}

@@

\section{Lessons Learned and Reflections}
\label{s:conclusion}

@@

In future work, it would be interesting to build statistical models of
the ``average'' programmer using each typing mode and compare their error
rates.
While one could start modeling with the \mnocheck{} and \mnonstrict{}
data, but there are too few \mstrict{} records at the moment due to our uniform
sampling rate.
More data is needed to experiment with models of \mstrict{}-mode programmers.

%% PS
%% - downloading from Kibana was a HUGE pain and source of errors. Awful!
%% - worth finding ways to reduce / ignore syntax errors?
